\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration

### 2016-03-30 paper/paper.tex
...

\section{Motivation}
In trajctory optimization with obstacles, it is often necessary to provide feasible inital trajectories. 
For example in robot motion planning, where the dimensions of the underlying spaces are rather high, this can be expensive. A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. An efficient approximative approach is using probabilistic roadmaps (PRMs).

@@

\subsection{Storage of the Graph}

- describe

- compare different models...?

@@

>>>>

\section{Motivation}
In trajectory optimization with obstacles, it is often necessary to provide feasible initial trajectories. 
For example in robot motion planning, where the dimensions of the underlying spaces are rather high, this can be expensive. A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. An efficient approximative approach is using probabilistic roadmaps (PRMs).

In this project, a parallelized version of the PRM method has been developed in general.
The robotics application then was implemented as an special case of this with GPU support.

@@

\subsection{Storing the Graph}

Apart from the indicator function evaluation, the main problem in building the probabilistic roadmap is, that for each new node $q$ it have to be determined all possible neighbors. That is the set of all nodes $\tilde q$ of the graph, with $\norm{\tilde q - q} \le D$.
If no special order of the nodes is given, one has to calculate the norms for all nodes of the graph, which is rather time consuming.

Therefore a special indexing is used here. 
The nodes are stored in a vector, which is organized in blocks with fixed sizes. The blocks are referred to by ids which are stored in a map.

\begin{lstlisting}
struct block{
	int pos;	 //! position of first vertex
	int num;	 //! current number of vertices stored
	block* next; //! if num==blocksize -> pointer to next block
};
\end{lstlisting}

\begin{lstlisting}
struct graph{
    std::map<int,block*> map; //map for accessing blocks
    std::vector<block> blocks;

    std::vector<float> qstorage; //length ndof*N

    int newblockpos;  //position of next new block
    int blocknum;     //number of used blocks
    .
    .
    .
};
\end{lstlisting}

The key of an block is computed by the mapping
\begin{equation*}
	key(q) = \lfloor q(1)/H \rfloor
\end{equation*}
such that $H>0$ becomes a parameter to adjust the grid density.
This has to be chosen so large that enough nodes are stored in one block in order to have efficient memory access.
On the other side, the smaller it is, the less computations have to be done for the neighborhood requests.
For illustration the following pseudocode shows the insertion of a node.

\begin{algorithm}[H]
	\KwData{$q\in\R^d$}
	Compute key=key(q)\\ 
	\eIf{key exists in map}{
		block=map[key]\\ 
		\While{ block full }{
			block=block->next
		}
	}{
		insert new block at map[key]
	}
	Insert node in block
	\caption{Inserting a node}
\end{algorithm}

@@

Tested have been made on a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
